A Tutorial on Tailored Simulation-Based Sample Size Planning for Experimental Designs with Generalized Linear Mixed Models

### 2024-06-21 manuscript.qmd
work on step 4

At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset to be collected in the planned study.
The next step is to specify how the statistical model shall be estimated in the actual study collected later.
This usually includes the selection of 1) a statistical framework (e.g., frequentist statistics), 2) a software package that is capable of estimating the model class of interest (e.g., the *lme4* R package), 3) an estimation algorithm (e.g., the default optimizer "bobyqa"), and 4) the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.

@@

In *lme4*, a GLMM is fitted with the `glmer` function. By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs. inaccurate (0) diagnosis.

@@

In the *lme4* output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.

>>>>

At this point, the researcher is capable of producing a simulated dataset similar to the actual dataset to be collected in the planned study.
The next step is to specify how the statistical model shall be estimated in the actual study collected later.
This requires the selection of a statistical framework, a software package that is capable of estimating the model class of interest, an estimation algorithm, and the specific model structure including all fixed effects, random effects, and the model family of the dependent variable.

@@

In *lme4*, a GLMM is fitted with the `glmer` function. By setting `family =  "binomial"`, we request a binomial GLMM appropriate for our binary dependent variable `y_bin` (the binomial GLMM uses the canonical logit link by default), which is defined as an accurate (1) vs. inaccurate (0) diagnosis. We use the default estimation algorithm (see `?glmerControl` for a list of alternative options).

@@

In the model output, the `Estimate` column in the `Fixed effects` table contains the estimates for the $\beta$ parameters, while the `Std.Dev.` column in the `Random effects` table contains the estimates for $\sigma_S$ and $\sigma_I$.
